---
title: "Reproducible Research - Peer Assessment 1"
output: html_document
---

####Loading and Preprocessing the Data

#####Load the Data
```{r}
setwd("~/RWD/RepDataPA1")
actdata <- read.csv("activity.csv", colClasses = c("numeric", "character", "numeric"))
head(actdata)
```

#####Process/Transform the data
```{r}
names(actdata)
```

```{r}
library(lattice)
```

```{r}
actdata$date <- as.Date(actdata$date, "%Y-%m-%d")
```



####What is Mean Total Number of Steps Per Day?


#####Calculate the total number of steps taken per day


```{r}
TotalSteps <- aggregate(steps ~ date, data = actdata, sum, na.rm = TRUE)
```

#####Plot a histogram of the total number of steps taken each day
```{r}
hist(TotalSteps$steps, main = "Total Steps By Day", xlab = "day", col = "blue")
```

#####Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(TotalSteps$steps)
median(TotalSteps$steps)
```

####What is The Average Daily Activity Pattern?

#####Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days

```{r}
time_series <- tapply(actdata$steps, actdata$interval, mean, na.rm = TRUE)
plot(row.names(time_series), time_series, type = "l", xlab = "5-min Interval", 
     ylab = "AVG Across All Days", main = "AVG Number of Steps Taken", 
     col = "blue")
```

#####Which 5-minute interval, on average across all of the days in the dataset, contains the maximum number of steps?

```{r}
IntervalMax <- which.max(time_series)
names(IntervalMax)
```

####Imputing Missing Values

#####Calculate and report the total number of missing values in the dataset

```{r}
activity_NA <- sum(is.na(actdata))
activity_NA
```

#####Devise a strategy for filling in all of the missing values in the dataset

```{r}
AvgSteps <- aggregate(steps ~ interval, data = actdata, FUN = mean)
NAData <- numeric()
for (i in 1:nrow(actdata)) {
        obs <- actdata[i, ]
        if (is.na(obs$steps)) {
                steps <- subset(AvgSteps, interval == obs$interval)$steps
        } else {
                steps <- obs$steps
        }
        NAData <- c(NAData, steps)
}
```

#####Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r}
new_actdata <- actdata
new_actdata$steps <- NAData
```

#####Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day.
######Histogram
```{r}
TotalSteps2 <- aggregate(steps ~ date, data = new_actdata, sum, na.rm = TRUE)
hist(TotalSteps2$steps, main = "Total steps by day", xlab = "day", col = "blue")
```

######Mean & Median:
```{r}
mean(TotalSteps2$steps)
median(TotalSteps2$steps)
```

######Results Analysis
The Mean has stayed the same but the Median has altered slightly. 

####Are There Differences In Activity Patterns Between Weekdays And Weekends?

#####Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
day <- weekdays(actdata$date)
daylevel <- vector()
for (i in 1:nrow(actdata)) {
        if (day[i] == "Saturday") {
                daylevel[i] <- "Weekend"
        } else if (day[i] == "Sunday") {
                daylevel[i] <- "Weekend"
        } else {
                daylevel[i] <- "Weekday"
        }
}
actdata$daylevel <- daylevel
actdata$daylevel <- factor(actdata$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = actdata, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")
```

#####Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

```{r}
xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number Of Steps")
```

